Constraint Grammar as a SAT problem

### 2015-05-18 paper/nodalida2015.tex
slightly more tofu

Applying logic to reductionist grammars has been explored earlier by \cite{lager98,lager_nivre01}, but it was never adopted for use.
% ; logic programming was too slow to be used for tagging or parsing. 
Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and 
they are used in domains such as microprocessor design and computational 
biology---these problems easily match or exceed CG in complexity. 
% Is this true for grammars with thousands of rules?

Due to these advances, we were able to revisit the idea and develop it further. 
% Whereas \cite{lager98} provides a proof-of-concept description, 
Missing from the previous logic-based works, 
we solve eventual rule conflicts by finding a solution that 
discards the least number of rules.
% allows the maximal number of rules to be applied.
We test our implementation by parsing texts in the order of 10,000s--100,000s
words, using grammars with hundreds of rules.

@@

%This increases the computational complexity, 
%but we are exploring ways in which it could simplify the grammar writing:
%our experiments show that less rules are needed, and features like ordering of the rules
%and requirement of being unambiguously tagged become irrelevant.

% Uncomment for full paper
% \section{Related work}
% \label{sect:related}

% Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
% and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
% \cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
% \cite{lager01transformation} discusses ordering of the CG rules.

@@

\paragraph{Rules disambiguate more.}
Considering our example phrase and rules, standard CG implementation
can only remove readings for the word which is tagged as \texttt{prn} or
\texttt{det}. The SAT-based implementation interprets the rules as
``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{1 n} or \texttt{1 v}).
This means that the rules are logically flipped: \texttt{REMOVE prn IF
  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
  IF (-1 prn)}. A rule with more conditions can be written as many
rules, each condition taking its turn to be in position to be removed
or selected. \\

@@

However, the results using our approach should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.
\pagebreak

>>>>

Applying logic to reductionist grammars has been explored earlier by \cite{lager98,lager_nivre01}, but it was never adopted for use.
% ; logic programming was too slow to be used for tagging or parsing. 
Since those works, SAT solving techniques have improved significantly \cite{marques_silva2010}, and they are used in domains such as microprocessor design and computational 
biology---these problems easily match or exceed CG in complexity. 
Due to these advances, we were able to revisit the idea and develop it
further.

Encoding CG in logic increases computational complexity, 
but as a tradeoff, it could simplify the grammar writing.
Due to representing the rules as implications, the requirement of being unambiguously tagged becomes irrelevant.
One rule corresponds to many logical formulas (each condition being
the target), making the number of the rules potentially smaller.
Ordering can be preserved or discarded; in the latter case, we solve
eventual rule conflicts by finding a solution that discards the least
number of rule applications.
% our experiments show that it is possible to write a small grammar, 
% which, when run with SAT-CG, disambiguates more than the same
%grammar run with VISL CG-3.

We test our implementation by parsing texts in the order of
10,000s--100,000s words, using grammars with hundreds of rules.

@@

% Uncomment for full paper
\section{Related work}
\label{sect:related}

Our work is inspired by \cite{lager98}, which presents constraint rules as a disjunctive logic program,
and \cite{lager_nivre01}, which reconstructs 4 different formalisms from a logical point of view.
\cite{lindberg_eineborg98ilp,asfrent14} use Inductive Logic Programming to learn CG rules from a tagged corpus.
\cite{lager01transformation} discusses ordering of the CG rules.

@@

\paragraph{Rules disambiguate more.}
Considering our example phrase and rules, standard CG implementation
can only remove readings for the word which is tagged as \texttt{prn} or
\texttt{det}. The SAT-based implementation interprets the rules as
``determiner and verb together are illegal'', and is free to take action that concerns also the word in the condition (\texttt{1 n} or \texttt{1 v}).
This means that the rules are logically flipped: \texttt{REMOVE prn IF
  (1 n)} translates into the same logical formula as  \texttt{REMOVE n
  IF (-1 prn)}. A rule with more conditions can be written as many
rules, each condition taking its turn to be the target of removal or selection. \\

@@

However, the results using our approach should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.
